---
title: "Random Forest"
author: "Philip Clark + Wyatt Pontius"
---

```{r}
library(randomForest)
library(ggplot2)
library(ROCR)
```

Reading in the data and separating out the dependent variable from the covariates.
```{r}
setwd("C:/Users/Wyatt/Documents/Year 3 - Autumn/MS&E 226/Project/Data/")
train_data <- read.csv(file = "final_train_data.csv")
validation_data <- read.csv(file = "final_validation_data.csv")
train_data[1] <- NULL
validation_data[1] <- NULL
dependent_variable_train <- as.factor(train_data[,15])
dependent_variable_validate <- as.factor(validation_data[,15])
independent_variable_train <- train_data[,-15]
independent_variable_validate <- validation_data[,-15]
```

A function which optimizes the value of mtry, the number of variables selected from the set of covariates when forming each tree
```{r}
tuneRF(independent_variable_train, dependent_variable_train)
```

Best mtry fluctuates around the default, so we'll run the model using the default mtry value (which is the square root of the number of covariates).
```{r}
warranty.rf <- randomForest(x = independent_variable_train, y = dependent_variable_train, ntree= 1000)
```

```{r}
print(warranty.rf)
varUsed(warranty.rf, by.tree = FALSE, count = TRUE)
```

A measure of the "importance" of each covariate, indicated by the increased in error accompanied by the removal of each covariate from the model
```{r}
importance(warranty.rf)
```

A plot of the error as the number of trees increases
```{r}
plot(warranty.rf)
```

Creating the prediction vector of the trained model across the training set.
```{r}
warranty.pred.train <- predict(warranty.rf, newdata = independent_variable_train)
ROCR.pred.train <- as.numeric(warranty.pred.train)
ROCR.actual.train <- as.numeric(dependent_variable_train)
pred.train = prediction(ROCR.pred.train, ROCR.actual.train)
```

Outputting the confusion matrix
```{r}
CM.train <- table(warranty.pred.train, dependent_variable_train)
print(CM.train)
```

Calculating values of the interest based on the confusion matrix
```{r}
accuracy.train <- sum(diag(CM.train))/sum(CM.train)
print(accuracy.train)
```

```{r}
precision.train <- CM.train[2,2]/(CM.train[2,2] + CM.train[2,1])
print(precision.train)
```

```{r}
recall.train <- CM.train[2,2]/(CM.train[2,2] + CM.train[1,2])
print(recall.train)
```

```{r}
perf_auc.train <- performance(pred.train, "auc")
auc.train <- perf_auc.train@y.values[[1]]
print(auc.train)
```

-------------------------------------------------------------------------------------------
VALIDATION DATA

```{r}
warranty.pred <- predict(warranty.rf, newdata = independent_variable_validate)
ROCR.pred <- as.numeric(warranty.pred)
ROCR.actual <- as.numeric(dependent_variable_validate)
pred = prediction(ROCR.pred, ROCR.actual)
```

```{r}
CM <- table(warranty.pred, dependent_variable_validate)
print(CM)
```

```{r}
accuracy <- sum(diag(CM))/sum(CM)
print(accuracy)
```

```{r}
precision <- CM[2,2]/(CM[2,2] + CM[2,1])
print(precision)
```

```{r}
recall <- CM[2,2]/(CM[2,2] + CM[1,2])
print(recall)
```

```{r}
perf_auc <- performance(pred, "auc")
auc <- perf_auc@y.values[[1]]
print(auc)
```
